Figure 2 and figure 2 supplementary material

### 2022-02-16 Figures/Figure2.Rmd
code organization and Sup.fig2.pdf generation
also did a change in the scatterplot axis labels and select and relocate
dplyr name conflicts

```{r}
# Does the typical deconvolution scatterplot comparing true proportions to certain deconvolution estimates. Called inside scatterplot_wrapper
scatterplot_truevsestimate <- function(merged_data, x, y, title, color){
    merged_data %>% dplyr::select(all_of(x),all_of(y)) %>%
    na.omit() -> to_plot

@@

  ## plot
  your_plot <- ggplot(to_plot) +
    aes_string(x = x, y = y) +
    geom_point(color = color) +
    labs(title = title, subtitle = stats, x = "", y = "") +
    geom_smooth(method=lm) +
    theme_classic() +
    theme(
      plot.title=element_text(family='', face='bold', colour='black', size=25, margin=margin(t=50,b=-40)),
      plot.subtitle=element_text(family='', face='italic', colour='black', size=18, margin=margin(t=50,b=-40)),
      text = element_text(size=15)
  )

@@

    ct_subset = data %>%
      dplyr::select(matches(ct)) %>%
      rename("H&E" = ct) %>%
      gather() %>%
      mutate(key = str_remove_all(key, paste(substring_remove, collapse = "|")),
             key=factor(key, levels=names(cor.ct.sorted)),
             method = as.factor(str_extract(key, "[^_]+")))
  adjust_x <- max(ct_subset$value)
  final_plot[[ct]] <- print(ggplot(ct_subset, aes(x = value, y = key, colour = method)) +
    geom_density_ridges2(jittered_points = TRUE, position = "raincloud", 
                         point_size = 0.25, point_alpha = 0.25, alpha = 0.7, rel_min_height = 0.005, point_color = "black") +
    theme_ridges() + ggtitle(ct) +
    annotate("text", x = adjust_x, y = 1:length(cor.ct.sorted), label = paste0("R = ",round(cor.ct.sorted, 2)), colour = "brown") +
    scale_x_continuous(limits = c(0, (adjust_x + adjust_x*0.2))) +
    scale_colour_manual(values = cols) +
     theme(axis.text.x = element_text(color = "black", size = 12, angle = 90, hjust = .5, vjust = .5, face = "plain"),
          axis.text.y = element_text(color = "black", size = 12, angle = 0, hjust = 1, vjust = 0, face = "plain"),  
          axis.title.x = element_text(color = "black", size = 10, angle = 0, hjust = .5, vjust = 0, face = "plain"),
          axis.title.y = element_text(color = "black", size = 10, angle = 90, hjust = .5, vjust = .5, face = "plain"),
          plot.title = element_text(color = "black", size = 20)) +
    ylab("") + xlab("% of cell type"))
  }
  return(final_plot)
}

@@

  identical(true_subset$ParticipantBarcode, deconv_subset_summary$Samples) %>% stopifnot()
  true_subset %>% bind_cols(deconv_subset_summary) %>% select(-c(Samples)) -> data_toplot

@@

  final_plot = list()
  color = "black"

  title = "ESTIMATE vs CCLE TIL10 Cancer"
  final_plot[["ccle"]] <- scatterplot_truevsestimate(merged_subset, "ESTIMATE", "Epidish_CCLE_TIL10_Cancer.cells", title, color)

  # Scatterplot BPRNACan
  title = "ESTIMATE vs BPRNACan Cancer"
  final_plot[["bprnacan"]] <- scatterplot_truevsestimate(merged_subset, "ESTIMATE", "Epidish_BPRNACan_cancer", title, color)

@@

summary_corrplots

@@

ridgeplots_by_cancer <- list()
tpurity_by_cancer <- list()
blank  <- ggplot() + theme_void()

for (cancer in unique(HE_all_TCGA$Study)){
  ridgeplots_by_cancer[[cancer]] <- wrapper_ridgeplot(deconv_data = tcga_deconv_ridge, true_estimates = HE_all_TCGA, cancer_abreviation = cancer)
  tpurity_by_cancer[[cancer]] <- tryCatch({ wrapper_tpurity(deconv_data = tcga_deconv, true_estimates = Aran_all_facs, cancer_abreviation = cancer) # sometimes there arent enough samples of the tumour type
        },
        error = function(cond) {
          message(paste("not enough FACs samples for ", cancer, "no scatterplot will be generated"))
          return(list(ccle = blank, bprna = blank))
        }
  )
  tpurity_by_cancer[[cancer]]$extra = blank
}

ridges <- ggarrange(plotlist = ridgeplots_by_cancer[["LUAD"]], ncol = 3, common.legend = T, legend = "bottom")
scater <- ggarrange(plotlist = tpurity_by_cancer[["LUAD"]], ncol = 3, common.legend = T)
blank  <- ggplot() + theme_void()
fullplot <- ggarrange(plotlist = list(scater, ridges, blank), nrow = 3, heights = c(0.15, 0.3, 0.45))
annotate_figure(fullplot, top = text_grob("LUAD", color = "black", size = 30))
pdf_name <- paste("test", "ridges", ".pdf", sep = "")
ggsave(pdf_name, dpi = "print", width = 210, height = 297, units = "mm", scale = 3)
# for (ct in c("Lymphocytes", "Macrophages", "Neutrophils")){
#    annotate_figure(print(ridgeplots_by_cancer[[cancer]][[ct]]),
#                              top = text_grob("Test", color = "black", size = 30),
#                              left = text_grob("Estimated fractions", color = "black", size = 20, rot = 90),
#                              bottom = text_grob("H&E estimates", color = "black", size = 20))
# }

#

@@

>>>>

```{r}
# Does the typical deconvolution scatterplot comparing true proportions to certain deconvolution estimates. Called inside scatterplot_wrapper
scatterplot_truevsestimate <- function(merged_data, x, y, title, color, x_lab = "", y_lab = ""){
    merged_data %>% dplyr::select(all_of(x),all_of(y)) %>%
    na.omit() -> to_plot

@@

  ## plot
  your_plot <- ggplot(to_plot) +
    aes_string(x = x, y = y) +
    geom_point(color = color) +
    labs(title = title, subtitle = stats, x = x_lab, y = y_lab) +
    geom_smooth(method=lm) +
    theme_classic() +
    theme(
      plot.title=element_text(family='', face='bold', colour='black', size=25, margin=margin(t=50,b=-40)),
      plot.subtitle=element_text(family='', face='italic', colour='black', size=18, margin=margin(t=50,b=-40)),
      text = element_text(size=15)
  )

@@

    ct_subset = data %>%
      dplyr::select(matches(ct)) %>%
      dplyr::rename("H&E" = ct) %>%
      gather() %>%
      mutate(key = str_remove_all(key, paste(substring_remove, collapse = "|")),
             key=factor(key, levels=names(cor.ct.sorted)),
             method = as.factor(str_extract(key, "[^_]+")))
  adjust_x <- max(ct_subset$value)
  final_plot[[ct]] <- print(ggplot(ct_subset, aes(x = value, y = key, colour = method)) +
    geom_density_ridges2(jittered_points = TRUE, position = "raincloud", 
                         point_size = 0.25, point_alpha = 0.25, alpha = 0.7, rel_min_height = 0.005, point_color = "black") +
    theme_ridges() + ggtitle(ct) +
    annotate("text", x = adjust_x, y = 1:length(cor.ct.sorted), label = paste0("R = ",round(cor.ct.sorted, 2)), colour = "brown") +
    scale_x_continuous(limits = c(0, (adjust_x + adjust_x*0.2))) +
    scale_colour_manual(values = cols) +
     theme(axis.text.x = element_text(color = "black", size = 12, angle = 90, hjust = .5, vjust = .5, face = "plain"),
          axis.text.y = element_text(color = "black", size = 12, angle = 0, hjust = 1, vjust = 0, face = "plain"),  
          axis.title.x = element_text(color = "black", size = 10, angle = 0, hjust = .5, vjust = 0, face = "plain"),
          axis.title.y = element_text(color = "black", size = 10, angle = 90, hjust = .5, vjust = .5, face = "plain"),
          plot.title = element_text(color = "black", size = 20)) +
    ylab("") + xlab("% of cell type"))
  }
  return(final_plot)
}

@@

  identical(true_subset$ParticipantBarcode, deconv_subset_summary$Samples) %>% stopifnot()
  true_subset %>% bind_cols(deconv_subset_summary) %>% dplyr::select(-c(Samples)) -> data_toplot

@@

  final_plot = list()
  color = "black"
  x_lab = "ESTIMATE"

  title = "ESTIMATE vs CCLE TIL10 Cancer"
  y_lab = "CCLE TIL10 Cancer"
  final_plot[["ccle"]] <- scatterplot_truevsestimate(merged_subset, "ESTIMATE", "Epidish_CCLE_TIL10_Cancer.cells", title, color,  x_lab, y_lab)

  # Scatterplot BPRNACan
  title = "ESTIMATE vs BPRNACan Cancer"
  y_lab = "BPRNACan Cancer"
  final_plot[["bprnacan"]] <- scatterplot_truevsestimate(merged_subset, "ESTIMATE", "Epidish_BPRNACan_cancer", title, color, x_lab, y_lab)

@@

@@

wrapper_ridgeplot(deconv_data = tcga_deconv_ridge, true_estimates = HE_all_TCGA, cancer_abreviation = "LUAD")
```

@@

```{r}
ridgeplots_by_cancer <- list()
tpurity_by_cancer <- list()
blank  <- ggplot() + theme_void()

for (cancer in unique(HE_all_TCGA$Study)){
  ridgeplots_by_cancer[[cancer]] <- wrapper_ridgeplot(deconv_data = tcga_deconv_ridge, true_estimates = HE_all_TCGA, cancer_abreviation = cancer)
  tpurity_by_cancer[[cancer]] <- tryCatch({ wrapper_tpurity(deconv_data = tcga_deconv, true_estimates = Aran_all_facs, cancer_abreviation = cancer) # sometimes there arent enough samples of the tumour type
        },
        error = function(cond) {
          message(paste("not enough FACs samples for ", cancer, "no scatterplot will be generated"))
          return(list(ccle = blank, bprna = blank))
        }
  )
  tpurity_by_cancer[[cancer]]$extra = blank

  ridges <- ggarrange(plotlist = ridgeplots_by_cancer[[cancer]], ncol = 3, common.legend = T, legend = "bottom")
  scater <- ggarrange(plotlist = tpurity_by_cancer[[cancer]], ncol = 3, common.legend = T)
  blank  <- ggplot() + theme_void()
  fullplot <- ggarrange(plotlist = list(scater, ridges, blank), nrow = 3, heights = c(0.15, 0.3, 0.45))
  annotate_figure(fullplot, top = text_grob(cancer, color = "black", size = 40))
  pdf_name <- paste("Sup.fig2_", cancer, ".pdf", sep = "")
  ggsave(pdf_name, dpi = "print", width = 210, height = 297, units = "mm", scale = 2.5)
}
system("pdfunite Sup.fig2_*.pdf Sup.fig2.pdf")
```
